---
title: "R hacks"
author: "Katerina Krizova"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  pdf_document: 
    toc: yes
    toc_depth: 3
---

* frequently used commands and hardly gained experience 
* R, R Markdown

# *WHAT NEEDS TO BE COVERED*

## apply, lapply, vapply aso

## read shapefiles: packages, quiet = T aso

## slice()

## grepl

# *INITIAL STEPS*

## PACKAGES

* tidyverse : 

```{r}
# install.packages("tidyverse")
require(tidyverse)
```

## DATA

* **starwars**
* **simple_table_4read.csv**
* **simple_table_2write.txt**

During this demonstration, We’ll use the *Starwars* characters dataset. 
The data, part of the dplyr package, come from [SWAPI](https://swapi.dev/), the Star Wars API. Since the data are integrated into *dplyr*, we don’t need to load the data, but you may still want to find information about the dataset. The codebood for the Starwars dataset is integrated into the dplyr documentation. To view the Starwars codebook, first load the library(dplyr) package. Then, via the Help pane, search starwars. Alternatively, in the Console pane, type ?starwars.

Note: dplyr is part of tidyverse which we load with the command: library(tidyverse)

```{r}
starwars
```

## SET PATHS

```{r}
path_in <- "D:/VURV/R_hacks/data/"
path_out <- "D:/VURV/R_hacks/results/"
```

# *R HACKS*

# READ AND WRITE

* read.table
* read.csv : sep is always ","

* write.table : usually set sep as ";"
* write.csv

## READ

header = T : 1st row is header

sep = "," OR ";" OR " "

fill = T : fills empty fields

read.table and read.csv is both fine

> absolute path

```{r}
df <- read.table("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
df <- read.csv("D:/VURV/R_hacks/data/simple_table_4read.csv", header = TRUE, sep = ";")
```

> relative paths

```{r}
df <- read.table(paste0(path_in, "simple_table_4read.csv"), header = TRUE, sep = ";")
df <- read.csv(paste0(path_in, "simple_table_4read.csv"), header = TRUE, sep = ";")
```

## WRITE

quote = F : string will be not written in quotes

row.names = F : no primary ID will be added

col.names = F : header won't be written

> absolute path

```{r}
write.table(df,"D:/VURV/R_hacks/results/simple_table_2write.txt", 
            sep = ";", quote = F, row.names = F) 
```

> relative paths

```{r}
write.table(df, paste0(path_out, "simple_table_2write.txt"), 
            sep = ";", quote = F, row.names = F) 
```


# FOLDERS, FILES AND LISTS

## create folder in direcory

```{r warning=F}
dir.create ("D:/VURV/R_hacks/results/new_folder_abs") # absolute path
dir.create (paste0(path_out, "new_folder_rel")) # relative path
```

## load all files from directory 

multiple files can be loaded from directory based on specific rule
mostly file type or name of the file

## names of files without extension

```{r}
library(tools)
file_path_sans_ext(list.files(pattern = "*.csv"))
```


## read files from direcory and combine them into 1 file

```{r}
acmnames <- list.files(paste0(path_out, crop, "/_outs/"), pattern= "\\.ACM$")  # Identify file names

acm_comb = data.frame() # empty dataframe to write files in

for(i in acmnames) { 
  a <- read.table(paste0(path_out, crop, "/_outs/", i), header = FALSE, skip = 11)
  a$source <- paste0(i)
  a$source <- str_sub(a$source,1,nchar(a$source)-4)
  df <- data.frame(a) 
  acm_comb <- rbind(acm_comb, df)
}
```


## split df based on a factor level

```{r eval=F}
sw_split <- split(starwars, starwars$homeworld) # creates list

lapply(names(sw_split), function(df) write.csv(sw_split[[df]], file=paste0(path_out, "new_folder_abs/", df, ".csv"), row.names = F)) 
```


## remove df from environment

remove single file or list of files with specific name pattern

```{r eval=F}
rm(df_round) # created later
rm(list=ls(pattern="sw_")) # created later
```


## delete folder

```{r}
for(i in 1:length(epic_parall)) { 
  unlink(paste0(path_epic, "EPIC_CS_", i, "/_outs/"), recursive = T) # If recursive = FALSE directories are not deleted, not even empty ones.
}
```


# NUMERICS

## round all

rounds all numeric values in DF

```{r}
df_round <- df %>% 
  mutate(across(where(is.numeric), ~ round(., digits = 1)))
```

## x digit numbers

creates numbers with specified number of digits

```{r}
# df %>% 
mutate(temp = sprintf("%04d", 1:1427))
```


# DATE

## date format

[Date Formats in R](https://www.r-bloggers.com/2013/08/date-formats-in-r/)

```{r}
df$sow_dat <- toString(df$sow_dat) # convert date into date format
df$sow_dat <- as.Date(df$sow_dat, "%Y%m%d") # convert date into date format

# doesn't work so far
# df$month <- format(df$sow_dat, "%m") # get months as numeric     
# df$year <- format(df$sow_dat,format="%y") # get years
```

## convert integer to date

```{r}
hvd_scen$IPLD <- as.Date(as.character(hvd_scen$IPLD), format = "%Y%m%d")
```

## date format with pipes

```{r}
df %>% 
mutate('11SPDatumZapisu' = format(as.Date(DatZap_S),"%d.%m.%Y")) 
```


# STRINGS

## starts with 

```{r eval=F}
sw_string <- starwars %>% 
  filter(startsWith(homeworld, "Tat"))
```

## remove last X characters

str_sub(species,**1**,nchar(species)**-2**))
1 -> new string strats with 1st character
-2 -> cut last 2 characters

```{r}
sw_rem <- starwars %>% 
  mutate(spec_short = str_sub(species,1,nchar(species)-2))
```


# SALT & PEPPER

```{r}
summary(starwars$species[starwars$species == "Human"])
summary(starwars$height[starwars$height > 100])
summary(starwars$height) # to compare with selected part of the table
```


# DATA ANALYSIS

## delete rows with NA values

```{r eval=F}
na.omit()
```

## case_when

creates a new column based on info from other columns
```{r}
sw_casewhen <- starwars %>% 
  mutate(power = case_when(sex == "female" ~ "female power", 
                           sex %in% c("male","none") ~ "too bad")) 
```

## if else

```{r}
#quest <- "nam_hom" # names and homeworld
quest <- "body" # name, heights, mass
#quest <- "color" # name, hair_color, skin_color, eye_color 
#quest <- "nope" # name, hair_color, skin_color, eye_color


if(quest == "nam_hom"){
     sw_quest <- starwars %>% 
    select(name, homeworld)
    print("Da wo ich herkomm")
  } else if(quest == "body"){
    sw_quest <- starwars %>% 
    select(name, height, mass)
    print("No one will jugdge you")
  } else if(quest == "color"){
    sw_quest <- starwars %>% 
    select(name, hair_color, skin_color, eye_color)
    print("You are all beautiful")
  } else{
    print("It's a trap!")
  }
```


# GGPLOT

## variables

## lab names

## lab scales

## legend name

labs(fill = "legend name")

## legend position

theme_void(legend.position = c(0.7, 0.2),legend.direction = "horizontal"))

## facet

### facet_grid

### facet_wrap

## ggsave

```{r eval=FALSE}
ggsave("D:/VURV/R_hacks/results/plot.png", device = "png", width = 10, height = 6, dpi = 300)
ggsave(paste0(path_out, "plot.png"), device = "png", width = 10, height = 6, dpi = 300)
ggsave(paste0(path_out, "plot.png"), device = "png", width = 10, height = 6, dpi = 300) # add variable to the name
```

# GIS 

## links to explore

http://oswaldosantos.github.io/ggsn/

## the best spatial analysis R packages

[source](https://towardsdatascience.com/the-best-spatial-analysis-packages-to-use-in-r-35855069f8b2)

* [*maps*](https://cran.r-project.org/web/packages/maps/maps.pdf)
    + basic package for producing simple maps very quickly and with little code
* [*spatstat*](https://cran.r-project.org/web/packages/spatstat/spatstat.pdf)
    + very good package to use on two dimensional spatial datasets
    + complete spatial statistics package that can be used for point pattern analysis
    + analyze spatial point pattern movement and distributions
    + Spatstat also allows you to conduct kernel density estimates (KDE)
* [*spdep*](https://cran.r-project.org/web/packages/spdep/spdep.pdf)
    + Spdep allows you to build spatial weights for a dataset and conduct autocorrelation on your dataset as well
* [*maptools*](https://cran.r-project.org/web/packages/maptools/maptools.pdf)
    + allows you to read in and manipulate spatial data
    + used for reading, writing, and exporting spatial data
    + From KML, SHP, and ASCII, if your data have coordinates you can use maptools to explore the information in it
* [*GISTools*](https://cran.r-project.org/web/packages/GISTools/GISTools.pdf)
    + Probably one of the most versatile and necessary GIS packages in R for spatial data
    + allows you to not only make maps but also to analyze data

##

### st_ functions

st_buffer(point, dist = 100)

## spatial ggplot

ggplot() +
  geom_path(aes(long, lat, group=group), data=worldUk, color="black", fill=NA) +
  coord_equal() +
  ggsn::scalebar(worldUk, dist = 100, st.size=3, height=0.01, dd2km = TRUE, model = 'WGS84')


# MARKDOWN

## handy links

* [Pimp my RMD](https://holtzy.github.io/Pimp-my-rmd/)

## document header

title: "Habits"
author: John Doe
date: March 22, 2005
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in


## r chunk options

* \`\`\` \{r\} \`\`\`
* r include = F : prevents from including the chunk in knitted file
* r echo = F : code not in file, results yes
* r eval = F : code is in the file, but the chunk does not run
* r message = F : does not show the message (e.i. "shp uploaded")
* r warning = F : does not show the warning (e.i. "folder already exists")

## headers

\# Title 1

\#\# Title 2

\#\#\# Title 3

\*text\*  *italics*

\*\*text\*\* **bold**

\~\~text\~\~ ~~struckthrough~~

## formatting

\\newpage braks the page

# FINAL INFO

[GIT](https://github.com/krizovakk/__R_hacks.git)