---
title: "GENERAL R HACKS by KK"
author: "Katerina Krizova"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, eval = F)
```
\newpage 

# GENERAL KAT NOTES

path_in = data
path_out = results

df = dataframe
level = level of a factor

read = simple table to read
write = simple file to save
f2s = file to save

L___ = list of files

floor = interval start
ceiling = interval end
number = operational number
string = operational string

# **INITIAL STEPS**

## Packages

```{r}
install.packages("tidyverse") # install
require(tidyverse) # "activate"
```

## Set paths

```{r}
path_in <- "D:/folder1/folder2/data/"
path_out <- "D:/folder1/folder2/results/"
```

### absolute path

```{r}
df <- read.table("D:/folder1/folder2/data/read.csv", header = TRUE, sep = ";")
df <- read.csv("D:/folder1/folder2/data/read.csv", header = TRUE, sep = ",")
```

### relative paths

```{r}
df <- read.table(paste0(path_in, "read.csv"), header = TRUE, sep = ";")
df <- read.csv(paste0(path_in, "read.csv"), header = TRUE, sep = ",")
```

## Environment

clear the entire environment

```{r}
rm(list=ls())
```

remove single file or list of files with specific name pattern

```{r eval=F}
rm(df) 
rm(df1, df2)
rm(list=ls(pattern="df_"))
```

remove all dataframes that contains "muz_" in their name

```{r}
rm(list = ls()[grepl("muz_", ls())])
```
\newpage

# **I M P O R T / E X P O R T**

df = dataframe
read = simple table to read
write = simple file to save

## TXT

read.table = [function](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
read.csv = [function](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
read_csv = [function]()

~ ENCODING ~
fileEncoding="cp1250"
encoding="utf-8"

### read txt

```{r}
df <- read.table("D:/VURV/R_hacks/data/read.txt", header = TRUE, sep = ";")
```

### read csv

read.csv : sep is always "," 

```{r}
df <- read.csv("D:/VURV/R_hacks/data/read.csv", header = TRUE, sep = ";")
```

### write to clipboard

```{r}
write.table(df, "clipboard", sep = "\t", row.names = FALSE)
```


### *memory issues*

\#Error in .jcall(cell, "V", "setCellValue", value) :  java.lang.OutOfMemoryError: GC overhead limit exceeded

```{r CHATGPT}

# Determine the number of rows and split point
total_rows <- nrow(APND_pk_rus)
split_points <- c(round(total_rows / 3), 2 * round(total_rows / 3))

# Split the dataframe into two parts
part1 <- APND_pk_rus[1:split_points[1], ]
part2 <- APND_pk_rus[(split_points[1] + 1):split_points[2], ]
part3 <- APND_pk_rus[(split_points[2] + 1):total_rows, ]

# Save each part to Excel files
gc()
write.xlsx(part1, file = paste0(path_ces, proj, "/", proj, "_PK_R_part1.xlsx"))
gc()
write.xlsx(part2, file = paste0(path_ces, proj, "/", proj, "_PK_R_part2.xlsx"))
gc()
write.xlsx(part3, file = paste0(path_ces, proj, "/", proj, "_PK_R_part3.xlsx"))

```

### conditional save

- zabrani prepsani jiz existujicich souboru

```{r}
f2s <- paste0(path_kk, proj, "_allXLSX.csv")

if(file.exists(f2s)){
  print("You've been here before.")
  rm(f2s)
} else {
  write.table(df, file = f2s, 
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")
  rm(f2s)
  print("File saved!")
  }
```

### factor save to multiple CSV

```{r}

df$level <- as.factor(df$level)

Ldf <- split(df, df$level)

for (i in seq_along(Ldf)) {
  subset_df <- Ldf[[i]]
  group_name <- unique(subset_df$level)
  # write.csv(subset_df, paste0(path_kk, "import/", proj, "_import_", group_name, ".csv"), row.names = FALSE)
  write.table(subset_df, file = "write.csv",
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="UTF-8")
}
```

## EXCEL

~ PACKAGES ~
 
 readxl, xlsx, openxlsx
 
~ FUNCTIONS ~
 
 readxl::read_excel()
 xlsx::read.xlsx()
 xlsx::write.xlsx()

### read excel

```{r}
df <- read_excel("D:/VURV/R_hacks/data/4read.xlsx" , sheet = 2, header = TRUE)
```

ISSUE: 'Warning: Expecting numeric in E1251 / R1251C5: got '41 (cm)'  
SOLUTION: df <- read_excel(paste0(path_snd, "muzbe_rozmery_ptaci.xlsx"), sheet = 1, **col_types="text"**)

### read multiple excel sheets

potrebny balik "data.table"

```{r}
read_excel_allsheets <- function(filename, tibble = FALSE) {
    # I prefer straight data.frames
    # but if you like tidyverse tibbles (the default with read_excel)
    # then just pass tibble = TRUE
    sheets <- readxl::excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X, col_types = "text"))
    if(!tibble) x <- lapply(x, as.data.frame)
    names(x) <- sheets
    x
}

mysheets <- read_excel_allsheets(paste0(path_proj, "20xx.xlsx")) # specifikuj excel file
# list_data <- Map(as.data.frame, mysheets) # nesjpis neni nutny krok
df <- rbindlist(mysheets) # pripadne list_data
```

### write excel

package: xlsx

```{r}
write.xlsx(x = df,                       
           file = "write.xlsx")
```

#### dataframes to sheets

- k existujicimu pripoji pomoci **append=TRUE**
- neprepisuje se. Pokud chci ulozit sheet, ktery uz v souboru je, hodi warning.

```{r}
a <- data.frame()
b <- data.frame()

write.xlsx(as.data.frame(a), file = "D:/folder1/folder2/write.xlsx", 
           sheetName = "dataframe A", row.names=FALSE)
write.xlsx(as.data.frame(b), file = "D:/folder1/folder2/write.xlsx",
           sheetName = "dataframe B", append=TRUE, row.names=FALSE)
```

#### levels to sheets

```{r}
install.packages("writexl")
library(writexl)

empty_data <- data.frame()
write_xlsx(empty_data, path = paste0(path_proj, "vystupy/", proj, "_PK_NV.xlsx")) 
  
for (i in Lzdroj) {
  subset_df <- save %>% filter(REL_zdroj == i)
  shtnam <- i
  write.xlsx(as.data.frame(subset_df), file = paste0(path_proj, "vystupy/", proj, "_PK_NV.xlsx"), 
             sheetName = i, row.names = FALSE, append = TRUE)
}
```

### *memory issue*

Error  
pouzit funkci gc() = Grabage Collection -> vycisti pamet po ulozeni predchoziho listu  

```{r}
exi <- paste0(path_ces, proj, "/", proj, "_KAT_PVO.xlsx")

if(file.exists(exi)){
  print("File already exists !!!")
} else {
  f2s = exi
  write.xlsx(as.data.frame(SAVE1), file= f2s, sheetName="sparovano", row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE2), file= f2s, sheetName="duplicity", append=T, row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE3), file= f2s, sheetName="nesparovanoMUS", append=T, row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE4), file= f2s, sheetName="nesparovanoCES", append=T, row.names=F, showNA = F)
  gc()
  write.xlsx(as.data.frame(SAVE0), file= f2s, sheetName="puv. vypis CES", append=T, row.names=F, showNA = F)
  rm(exi)
  rm(f2s)
  }
```

### save and style 

```{r}

wb <- loadWorkbook(paste0(path_in, "ICISHeren_harvestedReports_DA_WE_MA.xlsx"))
existing_data <- read.xlsx(wb, sheet = "Sheet1")
new_row <- df
start_row <- nrow(existing_data) + 2  # +2 to account for header row
writeData(wb, sheet = "Sheet1", x = new_row, 
          startRow = start_row, colNames = FALSE, na.string = "")

# style and save
style_DD <- createStyle(fgFill = "darkseagreen1")
style_TTF <- createStyle(fgFill = "powderblue")
style_THE <- createStyle(fgFill = "palegoldenrod")
style_VTP <- createStyle(fgFill = "mistyrose2")
addStyle(wb, sheet = "Sheet1", style = style_DD, cols = 3, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, sheet = "Sheet1", style = style_TTF, cols = 7:9, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, sheet = "Sheet1", style = style_TTF, cols = 21:23, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, sheet = "Sheet1", style = style_THE, cols = 11:13, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, sheet = "Sheet1", style = style_THE, cols = 25:27, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, sheet = "Sheet1", style = style_VTP, cols = 15:17, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, sheet = "Sheet1", style = style_VTP, cols = 29:31, rows = 1:1000, gridExpand = TRUE)
addStyle(wb, "Sheet1", createStyle(textDecoration = "bold"), rows = 1, cols = 1:100, gridExpand = TRUE)
freezePane(wb, sheet = "Sheet1", firstActiveRow = 2)
# setColWidths(wb, sheet = "Sheet1", cols = 1:100, widths = "auto") 
saveWorkbook(wb, paste0(path_out, "ICISHeren_harvestedReports_DA_WE_MA.xlsx"), overwrite = T)
```


## ACCESS

### connect db and read table

*.mdb = old MS ACCESS  
*.accdb = new MS ACCESS 

```{r}
require(RODBC) # package 

dbname <- paste0("---cesta+jmeno-databaze.accdb") 
con <- RODBC::odbcConnectAccess2007(dbname)      
RODBC::sqlTables(con)
tab <- RODBC::sqlFetch(con, "---jmeno-tabulky---")
```

### save to accdb

```{r}
# install.packages("odbc")
# install.packages("DBI")
library(odbc)
library(DBI)

con <- dbConnect(odbc::odbc(),
  .connection_string = paste0(
    "Driver={Microsoft Access Driver (*.mdb, *.accdb)};",
    "DBQ=", normalizePath("M:/03 klienti/narodni zemedelske muzeum - NZM/KK_workspace/test.accdb"), ";"))

dbWriteTable(con, "MyTable", osfo, overwrite = TRUE)

sapply(osfo, function(x) max(nchar(as.character(x)), na.rm = TRUE))

```


## SAMPLE

### sample (random)

```{r}
sam <- sample_n(df, 10) # n=10
```

### samplet top X rows

```{r}
sam10 <- df %>% 
  slice(1:10)
```
\newpage 

# **FOLDERS, FILES, LISTS**

## FOLDERS

###  create folder in direcory

```{r warning=F}
dir.create ("D:/folder1/folder2/results/new_folder_abs") # absolute path
dir.create (paste0(path_out, "new_folder_rel")) # relative path
```

### delete folder + content

## FILES

### load files from directory

multiple files can be loaded from directory based on specific rule
mostly file type or name of the file

### read multiple and combine 

```{r}
Ldf <- list.files(path_in, pattern= "\\.csv$")  # Identify file names

df_comb = data.frame() # empty dataframe to write files in

for(i in Ldf) { 
  a <- read.table(path_in, crop, i, header = FALSE, skip = 11) # skip = skip first 11 rows
  a$source <- paste0(i) #  enter source file name into a new column named "source"
  a$source <- str_sub(a$source,1,nchar(a$source)-4)
  df <- data.frame(a) 
  df_comb <- rbind(df_comb, df)
}
```

### split df based on a factor level

```{r eval=F}
df_split <- split(df, df$level) # creates list

lapply(names(df_split), function(df) write.csv(df_split[[df]], file=paste0(path_out, "new_folder_abs/", df, ".csv"), row.names = F)) 
```

## LISTS

### create list of files

```{r}
# one
Lvypis <- as.data.frame(list.files("D:/folder1/folder2/data/", full.names = T)) 

# more
Lvypis <- as.data.frame(list.files(c("D:/folder1/folder1/data/",
                                    "D:/folder1/folder2/data/",
                                    "D:/folder1/folder3/data/"),
                                  pattern = "jpg$|JPG$", full.names = T)) 

colnames(vypis)[1] <- "path"
```

### names of files without extension

```{r}
library(tools)
file_path_sans_ext(list.files(pattern = "*.csv"))
```
\newpage

# **DATA**

## NUMERICS

### round

round(x, #) = zaokrouhleni na # pocet decimalnich cisel
ceiling(x) = rounds values up to nearest integer
floor(x) = rounds values down to nearest integer
trunc(x) = truncates (cuts off) decimal places from values

```{r}
df_round <- df %>% 
  mutate(across(where(is.numeric), ~ round(., digits = 1)))
```

### X digit numbers

creates numbers with specified number of digits

```{r}
# df %>% 
mutate(temp = sprintf("%04d", 1:1427))
```

### sequence in interval

```{r}
mutate(int = mapply(function(floor, ceiling) paste(seq(floor, ceiling, by = 1), collapse = ";"), floor, ceiling))
```

### ID and sequence along

```{r PORADI}
v <- c("raz", "dva", "tři", "čtyři", "pět") 
v <- as.data.frame(v) %>% 
  mutate(por = seq_along(v),
         porplus = seq_along(v)+1)
```

### leading zeros -add

```{r LEADING ZEROS}
require(stringr) # package ::str_pad

new_number = str_pad(number, 6, pad = "0"))
```

mutate = create a new column  
new_number = name of a new column
str_pad = [function] (https://www.rdocumentation.org/packages/stringr/versions/1.5.0/topics/str_pad)
number = name of a source column (contains value '6' which we want add lead.zeros to)
6 = we want a final value to have exactly 6 places  
pad = "0" = empty places fill with "0"  

### leading zeros -remove

```{r}
new_number = sub("^0+", "", number)
```

## DATE

### check format

```{r}
class(date) # "Date"  # "character"
```

### date format

[Date Formats in R](https://www.r-bloggers.com/2013/08/date-formats-in-r/)

```{r}
df$date <- toString(df$date) # convert date into date format
df$date <- as.Date(df$date, "%Y%m%d") # convert date into date format

# doesn't work so far
# df$month <- format(df$sow_dat, "%m") # get months as numeric     
# df$year <- format(df$sow_dat,format="%y") # get years
```

### convert integer to date

```{r}
hvd_scen$IPLD <- as.Date(as.character(hvd_scen$IPLD), format = "%Y%m%d")
```

### date format with pipes

```{r}
df %>% 
mutate('11SPDatumZapisu' = format(as.Date(DatZap_S),"%d.%m.%Y")) 
```

### date conversion
   
28138 -> 13.01.1977
pokud mame oba formaty, resime pomoci vnorenych ifelsu takto:

```{r}
df <- source %>% 
  mutate(Mdat_ser = as.numeric(ifelse(str_detect(datumNabyti, "^.....$"), datumNabyti, NA)),
         Mdat_trans = format(as.Date(Mdat_ser, origin = "1899-12-30"), "%d.%m.%Y"),
         datnab22 = ifelse(str_detect(datumNabyti, "^.....$"), Mdat_trans, 
                           ifelse(str_detect(datumNabyti, "\\."), sub("\\.$", "", datumNabyti), datumNabyti)))
```

Mdat_ser = seriove datum
Mdat_trans = transformovane datum

### two date formats in one column

2025-04-12 and 15.06.2025 0:00

```{r}
egas2 <- egas %>% 
  mutate(DATE_parsed = case_when(
    grepl("^\\d{4}-\\d{2}-\\d{2}$", Date) ~ as.Date(Date, format = "%Y-%m-%d"),
    grepl("^\\d{2}\\.\\d{2}\\.\\d{4}", Date) ~ as.Date(Date, format = "%d.%m.%Y %H:%M"),
    TRUE ~ as.Date(NA)
  ))
```

## STRINGS

### length

```{r STRING LENGTH}
len = nchar(string)
```

### similarity

install.packages("RecordLinkage")
require("RecordLinkage")

levenshteinSim = funkce
column1, column2 = sloupce, ktere chci porovnat

```{r STRING SIMILARITY}
sim = levenshteinSim(column1, column2)
```

### detect case INsensitive

```{r STRING CASE SENS}
str_detect(column, regex("^FM\\s\\d+\\.jpg$", ignore_case = T))
```

### trim & lower case

```{r}
mutate(column2 = tolower(trimws(column1)))
```

### separate

```{r SPLIT STRING}

# # Split name column into firstname and last name
# df[c('First Name', 'Last Name')] <- str_split_fixed(df$Name, ' ', 2)
#  
# # Rearrange columns and remove original name column
# df <- df[c('First Name', 'Last Name', 'State')]

separate(trida,into=c("trida_CZ", "trida_LAT") , sep=" ")
separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge") # at first whitespace
separate(MDRUH, into=c("MROD", "rest") , sep=" ", extra = "merge", remove = F) # ponecha puvodni sloupec
```

### unite

creates a string of multiple column values with NAs.

```{r}
unite(new_column, c("kmen", "trida", "rad", "celed", "rod", "druh_Nomen"),
      sep = "/", na.rm = T, remove = FALSE) 
```

unite = [function](https://www.rdocumentation.org/packages/tidyr/versions/0.8.2/topics/unite)  
new_column = name of a new column  
c(***) = list of columns  
sep = sign for separation  
na.rm = skip columns containing NA  
remove = keep or delete source columns  

### remove last X characters

str_sub(species,**1**,nchar(species)**-2**))
1 -> new string strats with 1st character
-2 -> cut last 2 characters

```{r}
sw_rem <- starwars %>% 
  mutate(spec_short = str_sub(species,1,nchar(species)-2))
```


### seq_along with letters

```{r}
# lowercase
new_colum = case_when(n()!=1 ~ letters[seq_along(number)], TRUE ~ NA)
# uppercase
new_colum = case_when(n()!=1 ~ LETTERS[seq_along(number)], TRUE ~ NA)
```

### detect/extract/remove

functions:
  str_detect(column, pattern) = looking for a specific pattern
  str_extract(column, pattern) = extracts specific parts of a string
  str_remove(column, pattern) = deletes specific parts of a string

#### extract word

```{r}
mutate(boMROD = word(druh, 1)) # vezmi 1. slovo ze sloupce druh
```

\newpage

# **DATAFRAMES**

## EDA

### any NA ?

```{r}

```

### any \r or \n in text?

```{r}
any(grepl("[\r\n]", df))
```

### delete rows with NA values

```{r eval=F}
na.omit()
```

## General

### wide to long

```{r}
egsi <- df %>% 
  filter(index %in% c("CZ VTP", "THE", "TTF")) %>% 
  mutate('Delivery date' = date) %>% 
  pivot_wider(names_from = index, values_from = value)
```


### replace "" with NA

s pipe jsem zatim cestu nenasla

```{r}
taxon[taxon == ""] <- NA
```

### replace NA with ""

```{r}
ou <- taxon %>%
  replace(is.na(.), "")

ou <- taxon %>%
  mutate_all(~ ifelse(is.na(.), "", .))
```

### str_trim pro cely df

```{r}
df %>% 
  mutate(across(where(is.character), str_trim))
```

### mult column summary

```{r}
unique(df[,c('session','first','last')])

plyr::count(SAVE3[,c('zmena','duvod')]) # i s poctem
```

### split df based on a factor level

```{r eval=F}
sw_split <- split(starwars, starwars$homeworld) # creates list

lapply(names(sw_split), function(df) write.csv(sw_split[[df]], file=paste0(path_out, "new_folder_abs/", df, ".csv"), row.names = F)) 
```


### case_when

creates a new column based on info from other columns
```{r}
sw_casewhen <- starwars %>% 
  mutate(power = case_when(sex == "female" ~ "female power", 
                           sex %in% c("male","none") ~ "too bad")) 
```

## Columns

### remove columns acc to pattern

```{r}
bot <- bot %>% 
  select(-contains("_sort"), -nvg_id)
```

### remove last n columns

```{r}
select(-mus_invc, -(last_col(offset = 34):last_col()))
```

### remove multiple colums

typicky nazvy sloupcu z biolibu

```{r}
nam <- colnames(bl_sp)
noquote(nam)
nam2 <- paste0("-", nam)
noquote(nam2)
nam3 <- paste0(nam2, collapse = ",")
nam3
```

### add prefix to column names

```{r}
colnames(kmen) <- paste0("kmen_", colnames(kmen))
```

### rename columns based on pattern

```{r}
names(df) <- ifelse(grepl("pat1|pat2|pat3", names(df)), paste0("NEW NAME ", names(df)), names(df))
```


### merge two columns with not complete information

column 1 has some NAs  
column 2 has also some NAs but in different rows than column 1
TASK: create one column with merged info

coalesce = [function](https://www.rdocumentation.org/packages/dplyr/versions/1.0.10/topics/coalesce)

```{r}
#package
require(dplyr) # part of tidyverse

# command
mutate(kmen = coalesce(kmen_Nomen, trida_NadNomen))
```

### relocate columns

```{r}
relocate(MINVC, .before = JineC_S) 
relocate(MINVCDO, .after = MINVC)

# more than 1, in this order

relocate(Mrada, Mcislo,  Mcislo0, Msub_a, .after = prirc) 
```

### shift columns 

```{r}
rows_to_change <- a$zmena == "N" # filter rows to be shifted
a[rows_to_change,3:12] <- a[rows_to_change,2:12] # 3:12 nova lokace, 2:12 stara lokace
a[rows_to_change,2] <- NA # puvodni lukaci (diru) nahradit NA
```

## Rows

### DF top 10 rows

```{r}
mus_tab_cvic <- mus_tab %>% 
  slice(1:10)
```

### sample of 10 random rows

```{r}
ch <- sample_n(df, 10) # n=10

# random sample with selected columns
sample_n(aa, 10) %>% select(rok, prirc, porc104, subc105, prirc4)
```

### drop NA rows in spec col

```{r}
sez_int <- sez %>% 
  drop_na(MINVChorni)  
```

### filter with contains

```{r}
spo_ces <- ic %>% 
  filter(grepl("Další|Jiná|Lapidárium|Etnografická", zdroj))
```

### copy rows / multiply

1:nrow(df) = vsechny radky
, 1), ]) = jednou (tzn. ve vysledku dostaneme kazdy radek dvakrat)

```{r}
df <- rbind(df, df[rep(1:nrow(df), 1), ])
```

### duplicates

https://community.rstudio.com/t/filter-out-all-rows-with-duplicate-values/41043/4
```{r}
# Return names which have only a single row of data
data %>% 
  group_by(name) %>% 
  filter(n()==1)

# Return names which have more than one row of data
data %>% 
  group_by(name) %>% 
  filter(n() != 1)

# priznak
data %>% 
  group_by(name) %>% 
  mutate(priznak = ifelse(n()!=1), 1, 0)
```

### distinct

```{r}
distinct() # napric sloupci

distinct(sp_cislo, sp_id, .keep_all = TRUE) # vybrane sloupce, ponecha vse
```

### fill

```{r}
fill(Mrok, .direction = "down")
```

### drop row 

```{r}
df2 <- df1 %>% slice(-3) # based on row number
df2 <- df1 %>% slice(-which(condition = "something")) # based on condition
```

## Salt & pepper

```{r}
summary(starwars$species[starwars$species == "Human"])
summary(starwars$height[starwars$height > 100])
summary(starwars$height) # to compare with selected part of the table
```

## Save file with date

```{r}
write.table(img, file = paste0(path_data, "KK_upravyDat_proMigraci/prirustky/", format(Sys.Date(), "%Y%m%d"), "_KK_pokus.csv"), 
            quote = T, row.names = F, 
            sep = ";", dec = ",", 
            na = "", fileEncoding="cp1250")       
```

## Any \r or \n in text?

```{r}
any(grepl("[\r\n]", df))
```
\newpage

# **JOINS**

*inner_join* only keeps observations from x that have a matching key in y
*left_join()* keeps all observations in x.
*right_join()* keeps all observations in y.
*full_join()* keeps all observations in x and y.

[description and parameters](https://dplyr.tidyverse.org/reference/mutate-joins.html)

### Join only selected columns

```{r JOIN SEL COL}

a <- left_join(biolib %>% dplyr::select('4nazevKategorie', '5nomen', '8nadrizenyTaxon'),
            by = c("MNOMEN" = "5nomen"), keep = T)
```
\newpage

# **LOOPS**

## if, else if, else

Slouží ke kontrole podmínky a provedení různých bloků kódu podle výsledku (TRUE/FALSE).
```{r}
x <- 5

if (x > 0) {
  print("x je kladné")
} else if (x < 0) {
  print("x je záporné")
} else {
  print("x je nula")
}
```

## ifelse

Funkce pro rychlé podmíněné přiřazení na vektorové úrovni. Velmi užitečná při práci s datovými rámci (data frames).
```{r}
x <- c(1, -2, 3)
ifelse(x > 0, "kladné", "nekladné")
```

## for

Slouží k opakování výpočtů pro každou hodnotu ve vektoru či seznamu.
```{r}
for (i in 1:5) {
  print(paste("Hodnota i je", i))
}

```

## while 

Používá se, když předem nevíme, kolikrát se smyčka provede.
```{r}
i <- 1
while (i <= 5) {
  print(i)
  i <- i + 1
}
```

## repeat

Hodí se pro situace, kdy opakujeme, dokud nenastane určitá událost.
```{r}
i <- 1
repeat {
  print(i)
  i <- i + 1
  if (i > 5) {
    break
  }
}

```

## break, next

break ukončí smyčku předčasně
next přeskočí na další iteraci
```{r}
for (i in 1:5) {
  if (i == 3) next
  print(i)
}
```

## funkcionální alternativy k cyklům

Lepší (a často efektivnější) alternativy pro iteraci přes datové struktury v R:
  apply, lapply, sapply, tapply, map (z purrr) 

### apply()
– pro matice a data frames
```{r}
m <- matrix(1:9, nrow = 3)
apply(m, 1, sum)  # součet řádků
```

### lapply()
– seznam, výstup je seznam
```{r}
l <- list(a = 1:3, b = 4:6)
lapply(l, sum)
```

### sapply()
– zjednodušený výstup (např. vektor)
```{r}
sapply(l, sum)
```

### tapply()
– skupinove vypocty
```{r}
tapply(1:10, rep(c("A", "B"), each=5), mean)
```

### purrr::map()
– moderní funkce z tidyverse
```{r}
library(purrr)
map(l, sum)
```

## @@@ read all files from folder

1) make a list of files
```{r}
Lxls <- list.files(path_data, pattern = "*.xlsx")  # *.txt *.jpg ...
```

2) create empty dataframe and specify col namens
```{r}
df_filled <- data.frame()
# df_filled <- data.frame(matrix(ncol = 15, nrow = 0))

colnam <- c("prirc","rok_prir","dat_nabyt")
```
 
3) read files and bind them 
```{r}

for(i in Lxls) {
  a <- read_excel(paste0(path_data, i),
                  col_types = "text",
                  range = cell_cols("A:P"), na = " ") %>% # skip = 2 prepsano rangem
    slice(-1)
  names(a) <- colnam
  a$zdroj <- paste(i) # add source txt file name
  # print(count(a))
  df_filled <- rbind(df_filled, a)
  print(i)
  # rm(a)
}
```

\newpage

# **REGEX**

## Lookarounds

FOO BAR

**Positive Lookahead (?=pattern):**

Matches a group after the main expression if the group is followed by another expression.
Example: If you want to match "foo" only if it is followed by "bar", you can use the regex: foo(?=bar).

**Negative Lookahead (?!pattern):**

Matches a group after the main expression if the group is not followed by another expression.
Example: If you want to match "foo" only if it is not followed by "bar", you can use the regex: foo(?!bar).

**Positive Lookbehind (?<=pattern):**

Matches a group before the main expression if the group is preceded by another expression.
Example: If you want to match "foo" only if it is preceded by "bar", you can use the regex: (?<=bar)foo.

**Negative Lookbehind (?<!pattern):**

Matches a group before the main expression if the group is not preceded by another expression.
Example: If you want to match "foo" only if it is not preceded by "bar", you can use the regex: (?<!bar)foo.

## extract all numbers

```{r}

```

## extract all (except) letters

```{r}
str_detect(invc, "[A-Za-z]")
str_extract_all(invc, "[^a-zA-Z]")

```

## extract part ending with space or EOF

(?:\\s|$): This part matches either a whitespace character (\\s) or the end of the string ($). The (?: ... ) is a non-capturing group, which allows us to group the alternatives without capturing them as a separate match.

```{r}
extracted <- str_extract(text, "c\\.j\\..*?(?:\\s|$)")
```

## extract everything following "/"

```{r}
str_extract(predmet_cislo, "(?<=/).*$")
```

## replace more than one pattern

```{r}
str_replace_all(prirc, c("\\d+" = "", "\\/" = ""))
```

# **GGPLOT**

## Interesting sources

Various graph styles gallery [https://r-graph-gallery.com/]

## Basics

### dataframe for plots

Basic R has dataframe "mpg" integrated - could be used too

head(df)
  id      datum predikce1 predikce2
1  1 2025-04-12     25.02    24.959
2  2 2025-04-12     25.08    25.035
3  3 2025-04-12     25.10    25.048
4  4 2025-04-12     24.95    24.978
5  5 2025-04-12     24.76    24.801
6  6 2025-04-12     24.92    24.831


### create basic ggplot

Basics [https://ggplot2.tidyverse.org/reference/index.html]
- creates blank space for future plot
```{r}
ggplot(df)
```

### styles

geom_point [https://ggplot2.tidyverse.org/reference/geom_point.html]
geom_line [https://ggplot2.tidyverse.org/reference/geom_path.html]
geom_col [https://ggplot2.tidyverse.org/reference/geom_bar.html]

### variables

```{r}
ggplot(df)+
   geom_line(aes(x = id, y = predikce1))+
   geom_line(aes(x = id, y = predikce2))

# pro ucely ostatnich kapitol
g1 <- ggplot(df)+
   geom_line(aes(x = id, y = predikce1))

g2 <- ggplot(df)+
   geom_line(aes(x = id, y = predikce1))+
   geom_line(aes(x = id, y = predikce2))
```

### lab names

https://ggplot2.tidyverse.org/reference/labs.html

```{r}
g1 +
  labs(x = "ID", 
       y = "Predikce", 
       tag = "A",
       title = "Ilustracni graf", 
       subtitle = "ilustracni podnadpis", 
       caption = "(based on data from ...)") 
# colour = "predikce1" creates a legend

g1 +
  xlab("ID")+
  ylab("Predikce")+
  ggtitle("Ilustracni graf", subtitle = "ilustracni podnadpis")
```

### lab scales

```{r}
g1 +
  scale_x_continuous(name="ID", 
                     limits=c(800, 1500), 
                     breaks = c(1000,1200,1300),
                     labels = c("prvni", "druhy", "treti"))
```

### lab ticks

```{r}
g1 + 
  theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=14, angle=45))
g1 + theme_classic()+
  theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks = element_blank())
```


### legend name

labs(fill = "legend name")

### legend position

theme_void(legend.position = c(0.7, 0.2),legend.direction = "horizontal"))


### theme

Caution!
  theme() rewrites theme_classic()
  -> theme() as last element ALWAYS

```{r}
g1 + 
  theme_classic() # many others
```


### ggsave

```{r eval=FALSE}
ggsave("D:/VURV/R_hacks/results/plot.png", device = "png", width = 10, height = 6, dpi = 300)
ggsave(paste0(path_out, "plot.png"), device = "png", width = 10, height = 6, dpi = 300)
ggsave(paste0(path_out, "graf_energas_energodata_predikce_pro", sel_date), device = "jpg", width = 45, height = 9, dpi = 300) # add variable to the name
```

## Advanced

### trends

- zatim nefunguje -> doladit
```{r}
g1 +
  geom_smooth(formula = predikce1 ~ id, method = "lm")

```


### facet

#### facet_grid

#### facet_wrap

## Plotly

https://r-graph-gallery.com/plotly.html

```{r}
require(plotly)
library(htmlwidgets) # save the widget

a <- ggplotly(gvse)

saveWidget(a, file=paste0(path_out, "plotly-line-chart.html"))
```


# **GIS**

## links to explore

http://oswaldosantos.github.io/ggsn/

## the best spatial analysis R packages

[source](https://towardsdatascience.com/the-best-spatial-analysis-packages-to-use-in-r-35855069f8b2)

* [*maps*](https://cran.r-project.org/web/packages/maps/maps.pdf)
    + basic package for producing simple maps very quickly and with little code
* [*spatstat*](https://cran.r-project.org/web/packages/spatstat/spatstat.pdf)
    + very good package to use on two dimensional spatial datasets
    + complete spatial statistics package that can be used for point pattern analysis
    + analyze spatial point pattern movement and distributions
    + Spatstat also allows you to conduct kernel density estimates (KDE)
* [*spdep*](https://cran.r-project.org/web/packages/spdep/spdep.pdf)
    + Spdep allows you to build spatial weights for a dataset and conduct autocorrelation on your dataset as well
* [*maptools*](https://cran.r-project.org/web/packages/maptools/maptools.pdf)
    + allows you to read in and manipulate spatial data
    + used for reading, writing, and exporting spatial data
    + From KML, SHP, and ASCII, if your data have coordinates you can use maptools to explore the information in it
* [*GISTools*](https://cran.r-project.org/web/packages/GISTools/GISTools.pdf)
    + Probably one of the most versatile and necessary GIS packages in R for spatial data
    + allows you to not only make maps but also to analyze data

### st_ functions

st_buffer(point, dist = 100)

## spatial ggplot

ggplot() +
  geom_path(aes(long, lat, group=group), data=worldUk, color="black", fill=NA) +
  coord_equal() +
  ggsn::scalebar(worldUk, dist = 100, st.size=3, height=0.01, dd2km = TRUE, model = 'WGS84')


# MARKDOWN

## handy links

* [Pimp my RMD](https://holtzy.github.io/Pimp-my-rmd/)

## document header

title: "Habits"
author: John Doe
date: March 22, 2005
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in


## r chunk options

* \`\`\` \{r\} \`\`\`
* r include = F : prevents from including the chunk in knitted file
* r echo = F : code not in file, results yes
* r eval = F : code is in the file, but the chunk does not run
* r message = F : does not show the message (e.i. "shp uploaded")
* r warning = F : does not show the warning (e.i. "folder already exists")

## headers

\# Title 1

\#\# Title 2

\#\#\# Title 3

\*text\*  *italics*

\*\*text\*\* **bold**

\~\~text\~\~ ~~struckthrough~~

## formatting

\\newpage braks the page

# r e a l   t a s k s

## tvorba tabulky z vybraných sloupců

Je treba mit vytvoreny list policek, ktere maji byt ve vyslednem csv. Priklad: pro PK 'Lpk' obsahuje 105 nazvu poli pro CSV se 105 sloupci.
Zaroven je treba, aby se v pracovni tabulce nachazely prislusne sloupce se stejnym nazvem.
Funkce potom vyzobe a posklada vysledne CSV.

```{r}

# list sloupcu pro nove CSV

Lcol <- c("column1", "column2", "column3", "column4", "column5")

# vytvorit prazdny dataframe pro zapis

n <- 100 # pocet radku dle poctu zaznamu
df <- data.frame(smazme = 1:n) # prazdny dataframe; "smazme" sloupe s ID, pote smazat

# tvorba CSV

for(i in Lcol){
  if(i %in% colnames(source)){ # source = zdrojova tabulka
    ano <- source %>% select(all_of(i))
    df <- cbind(df, ano)
    print("jo")
  } else{
    ne <- data.frame(rep(NA, n))
    colnames(ne) <- i
    df <- cbind(df, ne)
    print("houby")}
}

df <- df %>% select(-smazme) 
```
