---
title: "Indexy plyn - ICIS Heren a EEX"
author: "Katerina Krizova"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in
---

# PATHS AND SETUP

```{r}
# packages
require(tidyverse)
require(openxlsx)
require(stringr)
require(writexl)

# paths
path_in <- "C:/Users/krizova/Documents/R/indexyPlyn/csv_oper/"
path_bu <- "C:/Users/krizova/Documents/R/indexyPlyn/d-1_backup/"
path_out <- "C:/Users/krizova/Documents/R/indexyPlyn/"

# development
# path_in <- "C:/Users/krizova/Documents/R/playground/indexyPlyn/"
# path_out <- "C:/Users/krizova/Documents/R/playground/indexyPlyn/"

# set local
Sys.setlocale("LC_TIME", "C")

# function
append_to_excel <- function(df, file_path, sheet) {
  library(openxlsx)
  wb <- loadWorkbook(file_path)  # Load existing workbook and read data from target sheet
  existing_data <- read.xlsx(wb, sheet = sheet) 
  next_row <- nrow(existing_data) + 2  # Calculate next empty row (account for header)
  writeData(  # Append new data to that row
    wb, 
    sheet = sheet, 
    x = df, 
    startRow = next_row, 
    colNames = FALSE,      # Don't write column names again
    na.string = ""
  )
  saveWorkbook(wb, file = file_path, overwrite = TRUE)  # Save workbook
}
```

# ICIS Heren

## Copy

```{r}
Ltrh <- c("TTF", "THE", "VTP") 
for (i in Ltrh) {
  files <- list.files("X:/ICIS_Heren/", pattern = paste0(i, ".csv"), full.names = T) 
  if (length(files) > 0) {
    info <- file.info(files)
    newest_file <- rownames(info)[which.max(info$mtime)]
    file.copy(from = newest_file,
          to = path_in)
  }
} 

rm(info)
```

## Read and Process

ttf 35
the 34
cz  11
cons 35

```{r}
Ltrh <- c("TTF", "THE", "VTP")
for (i in Ltrh) {
  files <- list.files(path_in, pattern = paste0(i, ".csv"), full.names = T) 
  if (length(files) > 0) {
    info <- file.info(files)
    newest_file <- rownames(info)[which.max(info$mtime)]
    df <- read.csv(newest_file, header = TRUE, sep = ",") %>% 
      filter(str_detect(Name, "^Natural Gas")) %>% 
      mutate(publ = as.Date(Date, format = "%d %b %Y"),
             weekday = weekdays(publ),
             Den = case_when(weekday == "Monday" ~ "po",
                             weekday == "Tuesday" ~ "ut",
                             weekday == "Wednesday" ~ "st",
                             weekday == "Thursday" ~ "ct",
                             weekday == "Friday" ~ "pa",
                             weekday == "Saturday" ~ "so",
                             weekday == "Sunday" ~ "ne"),
             Indicator = str_remove(Name, " TTF| THE| Czech Republic"),
             deli = case_when(Den == "pa" & Indicator == "Natural Gas Day-ahead" ~ publ  + 3, 
                              Den == "pa" & Indicator == "Natural Gas Weekend+0" ~ publ  + 1,
                              TRUE ~ publ + 1),
             deli_day = case_when(Den == "pa" & Indicator == "Natural Gas Day-ahead" ~ "po", 
                                  Den == "po" & Indicator == "Natural Gas Day-ahead" ~ "ut",
                                  Den == "ut" & Indicator == "Natural Gas Day-ahead" ~ "st",
                                  Den == "st" & Indicator == "Natural Gas Day-ahead" ~ "ct",
                                  Den == "ct" & Indicator == "Natural Gas Day-ahead" ~ "pa",
                                  Den == "pa" & Indicator == "Natural Gas Weekend+0" ~ "so+ne",
                                  TRUE ~ NA))
    print(newest_file)
    rm(info)
    rm(newest_file)
  } else {
    warning("Not the robots you're looking for.")
  }
  if (i == "TTF"){
    df <- df %>% select("Publication day" = Den, publ, deli, "Delivery day" = deli_day, Indicator, Period, "TTF Bid" = Bid, "TTF Offer" = Offer, "TTF Mid" = Mid.Price)
    assign("ttf", df)
  } else if (i == "THE"){
    df <- df %>% select(Indicator, "THE Bid" = Bid, "THE Offer" = Offer, "THE Mid" = Mid.Price) 
    assign("the", df)
  } else if (i == "VTP"){
    df <- df %>% select(Indicator, "CZ-VTP Bid" = Bid, "CZ-VTP Offer" = Offer, "CZ-VTP Mid" = Mid.Price) 
    assign("vtp", df)
  } else {
    warning("I have a bad feeling about this.")
  }
  rm(df)
  rm(i)
  rm(files)
} 

cons <- left_join(ttf, the, by = "Indicator") %>%
  left_join(vtp, by = "Indicator") %>%
  mutate(across(c('TTF Bid', 'TTF Offer', 'TTF Mid',
                  'THE Bid', 'THE Offer', 'THE Mid',
                  'CZ-VTP Bid', 'CZ-VTP Offer', 'CZ-VTP Mid'), as.numeric)) %>% 
   mutate(across(where(is.character), ~ replace(., is.na(.), "")))
```
### RESULT 1: DA, WE, MA

```{r}
da <- cons %>% 
  filter(ifelse(`Publication day` == "pa", Indicator %in% c("Natural Gas Weekend+0", "Natural Gas Day-ahead"), Indicator == "Natural Gas Day-ahead")) %>% 
  filter(str_detect(Indicator, "^Natural"))
names(da) <- ifelse(grepl("Bid|Offer|Mid", names(da)), paste0("(d) ", names(da)), names(da))

mo <- cons %>% 
  filter(str_detect(Indicator, "Natural Gas Month.1")) %>%  # z nejakeho duvodu chce cely nazev. regex nefunguje
  select(-'Publication day', -deli, -publ, -'Delivery day')
names(mo) <- ifelse(grepl("Indicator|Period|Bid|Offer|Mid", names(mo)), paste0("(m) ", names(mo)), names(mo))

lines <- cbind(da, mo) 
day <- unique(lines$`Publication day`)

if (day == "pa"){
  lines <- rbind(lines, lines[rep(2, 1), ])
  # lines$deli[lines$Indicator == "Natural Gas Day-ahead"] <- lines$publ+3
  lines$deli[nrow(lines)] <- lines$publ+2
  lines <- lines[order(lines$deli), ]
} else{
  lines$deli <- lines$publ+1
}

dawemo <- lines %>% 
  mutate('Publication date' = format(publ, "%d.%m.%Y"),
         'Delivery date' = format(deli, "%d.%m.%Y")) %>% 
  select(-publ, -deli) %>% 
  relocate('Publication date', 'Delivery date', .after = `Publication day`)

dawemo$`Publication date` <- as.Date(dawemo$`Publication date`, format = "%d.%m.%Y")
dawemo$`Delivery date` <- as.Date(dawemo$`Delivery date`, format = "%d.%m.%Y")

# vymazat vikend pro MA !
blank <- c("(m) TTF Bid", "(m) TTF Offer", "(m) TTF Mid", "(m) THE Bid", "(m) THE Offer", "(m) THE Mid", "(m) CZ-VTP Bid","(m) CZ-VTP Offer", "(m) CZ-VTP Mid")
dawemo[2:3, blank] <- ""
dawemo[blank] <- sapply(dawemo[blank],as.numeric)

rm(lines)

# append
append_to_excel(dawemo, paste0(path_out, "indexyPlyn_AKTUALNI.xlsx"), sheet = "ICIS")
```

here we get a warning - checked, should be fine. 

### RESULT 2: COMPLETE ICIS

```{r}
icis <- cons %>% 
 mutate(`THE/TTF` = `THE Mid`-`TTF Mid`,
         `CZ-VTP/TTF` = `CZ-VTP Mid`-`TTF Mid`,
         `CZ-VTP/THE` = `CZ-VTP Mid`-`THE Mid`,
         `Publication date` = format(publ, "%d.%m.%Y"),
         `Delivery date` = ifelse(str_detect(Indicator, "Day-ahead"), format(deli, "%d.%m.%Y"), "")) %>% 
  select(-publ, -deli) %>% 
  relocate(`Publication date`, `Delivery date`, .after = `Publication day`)

append_to_excel(icis, paste0(path_out, "indexyPlyn_AKTUALNI.xlsx"), sheet = "ICIS komplet")
```

## Back-up

-- optional
-- execute when satisfied with the result
-- creates back-up copy for the day

# EEX

```{r}

# PO
# pub_d <- as.Date("29.8.2025", format = "%d.%m.%Y") # zadej patek
# del_d <- as.Date(pub_d+1)

# UT-PA
del_d <- as.Date(Sys.Date(), format = "%d.%m.%Y")
pub_d <- as.Date(Sys.Date()-1, format = "%d.%m.%Y")

weekday = weekdays(pub_d)
```

## EOD

segment <tbody>
specifikace: text-align: right

WEEKEND:

resen automaticky: DA v patek = vikend -> sam se namnozi na tri radky -> treba pak upravit pondeli - vychazi v nedeli, prepsat rucne, nema smysl automatizovat

```{r}
# read from clipboard
text <- paste(readClipboard(), collapse = "\n")
# print(text)

# match indices and prices

pattern <- '<tr class="mv-quote-row".*?><td style="text-align: left;">(.*?)</td><td style="text-align: right;">(.*?)</td><td style="text-align: right;">(.*?)</td><td style="text-align: right;">(.*?)</td>'

matches <- str_match_all(text, pattern)[[1]]

# Create data frame with raw string values preserved
odf <- data.frame(
  index = matches[, 2],
  lastPrice = str_trim(matches[, 3]),
  lastVolume = str_trim(matches[, 4]),
  EOD = str_trim(matches[, 5]),
  stringsAsFactors = FALSE
) %>% 
  mutate(lastPrice = str_remove(lastPrice, "-"), 
         lastVolume = str_remove(lastVolume, "&nbsp;"),
         EOD = as.numeric(str_replace(EOD, ",", "\\."))) %>% 
  select(index, EOD)

if (weekday == "Friday"){
  we <- odf %>% filter(index %in% c("CZ VTP Weekend", "THE Weekend", "TTF Weekend"))  %>% 
    mutate(market = paste("EOD:", str_extract(index, "CZ VTP|THE|TTF")),
           'Delivery date' = del_d,
           Index = paste0("EEX Spot EOD")) %>%
    select(-index, -Index) %>%
    pivot_wider(names_from = market, values_from = EOD)
  da <- odf %>% filter(index %in% c("CZ VTP Day Ahead", "THE Day Ahead", "TTF Day Ahead")) %>% 
    mutate(market = paste("EOD:", str_extract(index, "CZ VTP|THE|TTF")),
           'Delivery date' = pub_d+3,
           Index = paste0("EEX Spot EOD")) %>%
    select(-index, -Index) %>%
    pivot_wider(names_from = market, values_from = EOD)
  eod <- rbind(we, we[rep(1:nrow(we), 1), ], da)
  eod[2, 'Delivery date'] <- del_d+1
} else {
  eod <- odf %>% filter(index %in% c("CZ VTP Day Ahead", "THE Day Ahead", "TTF Day Ahead")) %>% 
    mutate(market =  paste("EOD:", str_extract(index, "CZ VTP|THE|TTF")),
           'Delivery date' = del_d,
           Index = paste0("EEX Spot EOD")) %>%
    select(-index, -Index) %>%
    pivot_wider(names_from = market, values_from = EOD)
}
```

## EGSI

DAY a WEEKEND jsou v rozdilnych tabulkach -> kopirujeme to, co je relevantni

segment <tbody>
specifikace: text-align: left

```{r}
# read from clipboard
text <- readClipboard()
# print(text)

# match indices and prices
matches <- str_match_all(
  text,
  '<td style="text-align: left;">(.*?)</td><td style="text-align: left;">(\\d{1,3}(?:,\\d{3})*)</td>'
)[[1]]

# create table
gdf <- data.frame(
  market = matches[, 2],
  value = matches[, 3],
  stringsAsFactors = FALSE
) %>%
  mutate(value = round(as.numeric(str_replace_all(value, ",", "")), 3))  # remove commas

gdf <- data.frame(
  market = matches[, 2],
  value = matches[, 3],
  stringsAsFactors = FALSE
) %>% 
  mutate(value = round(as.numeric(str_replace(value, ",", ".")), 3))

egsi <- gdf %>% 
  filter(market %in% c("CZ VTP", "THE", "TTF")) %>% 
  mutate(market = paste("EGSI:", market)) %>% 
  pivot_wider(names_from = market, values_from = value) 
```

# FUTURES

nelze automatizovat (prilis mnoho klikani pro volbu trhu a produktu)
vyplnuejme jen 99999, nutno prepsat rucne v excelu nebo zde

```{r}
emp <- data.frame(
  date = character(),
  value1 = numeric(),
  value2 = numeric(),
  value3 = numeric(),
  value4 = numeric(),
  value5 = numeric(),
  value6 = numeric(),
  value7 = numeric(),
  value8 = numeric(),
  value9 = numeric(),
  stringsAsFactors = FALSE
)

# PO
# df <- data.frame(
#   `Publication date` = pub_d,
#   value1 = 99.999, # VTP Cal+1
#   value2 = 99.999, # VTP Q+1
#   value3 = 99.999, # VTP M+1
#   value4 = 99.999, # THE Cal+1
#   value5 = 99.999, # THE Q+1
#   value6 = 99.999, # THE M+1
#   value7 = 99.999, # TTF Cal+1
#   value8 = 99.999, # TTF Q+1
#   value9 = 99.999  # TTF M+1
# )

# UT-PA
df <- data.frame(
  `Publication date` = pub_d,
  value1 = 34.112, # VTP Cal+1
  value2 = 34.570, # VTP Q+1
  value3 = 33.600	, # VTP M+1
  value4 = 32.978, # THE Cal+1
  value5 = 33.591, # THE Q+1
  value6 = 32.747, # THE M+1
  value7 = 31.656, # TTF Cal+1
  value8 = 32.467	, # TTF Q+1
  value9 = 31.769		  # TTF M+1
)

futu <- rbind(emp, df)
names(futu) <- c("Publication date", "CZ VTP Cal+1", "CZ VTP Q+1", "CZ VTP M+1",
                    "THE Cal+1", "THE Q+1", "THE M+1",
                    "TTF Cal+1", "TTF Q+1", "TTF M+1")
```

# CBIND AND APPEND

```{r}
eex <- cbind(eod, egsi, futu)
append_to_excel(eex, paste0(path_out, "indexyPlyn_AKTUALNI.xlsx"), sheet = "EEX")
```

# F I N A L  SAVE AND BACKUP

```{r}
# DELETE OPER CSV ==============================================================

unlink(paste0(path_in, "*"), recursive = FALSE) # smaze vse !!!

# PERSONAL BACKUP COPY =================== uncomment ===========================

# file.copy(from = paste0(path_out, "indexyPlyn_AKTUALNI.xlsx"),
#           to = path_bu,
#           overwrite = TRUE)
# 
# file.copy(from = paste0(path_out, "indexyPlyn_AKTUALNI.xlsx"),
#           to = "X:/Nakup _ NEW/Indexy_Plyn/indexyPlyn_AKTUALNI.xlsx",
#           overwrite = TRUE)

# CLEAR ENVIRONMENT

# rm(list=ls())
```
# PLOTS

## 
